Aa Language Reference Manual

### 2016-10-23 v2/Ahir-asm/devel/docs/AaLRM.tex
Added parameter support to be used to parametrize an Aa program to a certain extent.

Variables in  the program can belong to the following classes:
\begin{itemize}
\item {\bf Parameters} can be used to configure the Aa program to
a certain extent.  These are 32-bit integers declared as
\begin{verbatim}
$parameter PIPEDEPTH 16
$parameter PIPEDEPTH2  (2 * PIPEDEPTH) 
\end{verbatim}
for example.  
Parameters can be used to specify some 
constants inside
the program (this will be indicated whenever relevant).
Parameters
should be known at the point at which they are used.  Also,
parameters can be declared only at global program scope (that is
not inside a module).
\item {\bf Storage} variables declared in the program
can be written into from multiple
points in the program.  When declared at the program
scope, storage variables are visible throughout
the program.  Storage variables can be thought of 
as memory locations which can be accessed from different
points in the program.
A storage variable declaration has the form
\begin{verbatim}
$storage <var-name>: <var-type> 
\end{verbatim}
For example
\begin{verbatim}
$storage mempool: $array [1024] $of $uint<32>
\end{verbatim}
\item {\bf Pipe} variables declared at the program scope
are globally visible buffers (which
can be defined to be either first-in-first-out (FIFO) or
last-in-first-out (LIFO)) that can 
be read from anywhere in the program and can be
written into from anywhere in th program.  A pipe
is a FIFO (or FIFO) buffer with a specified number of
slots.  A write to a pipe succeeds
only if there is an empty slot in the FIFO (a successful
write uses up an empty slots.  A read
from the pipe succeeds only if there is at
least one occupied slot in the pipe (a successful read
creates an empty slot in the pipe). 
Thus, pipes can be used to model queues as
well as stacks, and also provide a mechanism for synchronization and
for message passing in an \Aa program.
\begin{itemize}
\item A pipe that is only written into in the program, but
is not read from, is assumed to have a destination outside
the program.  Such a pipe is an output port of the
program.
\item A pipe that is only read from in a program but not
written into, is assumed to have a source outside the program.
Such a pipe is an input port of the program.
\end{itemize}
A pipe declaration has the form
\begin{verbatim}
($lifo | $noblock)? $pipe <pipe-name>: <pipe-type>  ($depth <integer>)?
\end{verbatim}
For example
\begin{verbatim}
$pipe myfifo: $uint<32> $depth 16 // 16-slot queue
$lifo $pipe mystack: $uint<32> $depth 4 // 4-slot stack
$noblock $pipe cmd_pipe: $uint<8> // non-blocking reads.
\end{verbatim}
A non-blocking pipe when read, will deliver $0$ if
it does not contain data (thus, you will have to reserve
the $0$ for invalid data).

@@

Constants can be specified in one of many
ways, depending on their type.
\begin{itemize}
\item Integers can be specified in their
decimal form, as for example 
\begin{verbatim}
23
-8
\end{verbatim}
or in binary form, as for example
\begin{verbatim}
_b10111
\end{verbatim}
When specified in binary form, the twos-complement
value of the integer being specified should be
used.
\item Floats are specified in the exponentiated
form
\begin{verbatim}
_f2.3000e+10
_f-1.354e+10
\end{verbatim}
where the mantissa has to have exactly one digit
to the left of the decimal point.
\item Composite constants are specified
as a space separated list of values.
\begin{verbatim}
( 12 32 43 10)
( (1 5) (2 8) (9 100) )
\end{verbatim}
etc.  The elements are listed in 
row major form.  Thus a[2][2] is
listed as
\begin{verbatim}
a[0][0] a[0][1] a[1][0] a[1][1]
\end{verbatim}
\end{itemize}

@@

>>>>

Variables in  the program can belong to the following classes:
\begin{itemize}
\item {\bf Parameters} can be used to configure the Aa program to
a certain extent.  These are 32-bit integers declared as
\begin{verbatim}
$parameter PIPEDEPTH 16
$parameter PIPEDEPTH2  {2 * PIPEDEPTH} 
\end{verbatim}
for example.    Note that the parameter
PIPEDEPTH2 is obtained by evaluating the
parameter expression 
\begin{verbatim}
{2 * PIPEDEPTH} 
\end{verbatim}
Parameter expressions  are described in more detail Section \ref{sec:ParamExprs}.
Parameters can be used to specify some 
constants inside
the program (this will be indicated whenever relevant).
Parameters
should be known at the point at which they are used.  Also,
parameters can be declared only at global program scope (that is
not inside a module).
\item {\bf Storage} variables declared in the program
can be written into from multiple
points in the program.  When declared at the program
scope, storage variables are visible throughout
the program.  Storage variables can be thought of 
as memory locations which can be accessed from different
points in the program.
A storage variable declaration has the form
\begin{verbatim}
$storage <var-name>: <var-type> 
\end{verbatim}
For example
\begin{verbatim}
$storage mempool: $array [1024] $of $uint<32>
\end{verbatim}
\item {\bf Pipe} variables declared at the program scope
are globally visible buffers (which
can be defined to be either first-in-first-out (FIFO) or
last-in-first-out (LIFO)) that can 
be read from anywhere in the program and can be
written into from anywhere in th program.  A pipe
is a FIFO (or FIFO) buffer with a specified number of
slots.  A write to a pipe succeeds
only if there is an empty slot in the FIFO (a successful
write uses up an empty slots.  A read
from the pipe succeeds only if there is at
least one occupied slot in the pipe (a successful read
creates an empty slot in the pipe). 
Thus, pipes can be used to model queues as
well as stacks, and also provide a mechanism for synchronization and
for message passing in an \Aa program.
\begin{itemize}
\item A pipe that is only written into in the program, but
is not read from, is assumed to have a destination outside
the program.  Such a pipe is an output port of the
program.
\item A pipe that is only read from in a program but not
written into, is assumed to have a source outside the program.
Such a pipe is an input port of the program.
\end{itemize}
A pipe declaration has the form
\begin{verbatim}
($lifo | $noblock)? $pipe <pipe-name>: <pipe-type>  ($depth <integer>)?
\end{verbatim}
For example
\begin{verbatim}
$pipe myfifo: $uint<32> $depth 16 // 16-slot queue
$lifo $pipe mystack: $uint<32> $depth 4 // 4-slot stack
$noblock $pipe cmd_pipe: $uint<8> // non-blocking reads.
\end{verbatim}
A non-blocking pipe when read, will deliver $0$ if
it does not contain data (thus, you will have to reserve
the $0$ for invalid data).

@@

Constants can be specified in one of many
ways, depending on their type.
\begin{itemize}
\item Integers can be specified in their
decimal form, as for example 
\begin{verbatim}
23
-8
\end{verbatim}
or in binary form, as for example
\begin{verbatim}
_b10111
\end{verbatim}
When specified in binary form, the twos-complement
value of the integer being specified should be
used.
\item Floats are specified in the exponentiated
form
\begin{verbatim}
_f2.3000e+10
_f-1.354e+10
\end{verbatim}
where the mantissa has to have exactly one digit
to the left of the decimal point.
\item Composite constants are specified
as a space separated list of values.
\begin{verbatim}
( 12 32 43 10)
( (1 5) (2 8) (9 100) )
\end{verbatim}
etc.  The elements are listed in 
row major form.  Thus a[2][2] is
listed as
\begin{verbatim}
a[0][0] a[0][1] a[1][0] a[1][1]
\end{verbatim}
\item Parameters can be used as simple identifiers.
These are converted to integers (whose width as a uint
is unknown) provided they are defined at the point
of reference.
\end{itemize}

@@

\section{Parameter Expressions} \label{sec:ParamExprs}

Parameters are like constants, but have an implicit
integer types.  They are to be used in lieu of
constants in order to provide a measure of configurability
to the Aa program, as well as to improve readability of
the code.

Pararmeters can be combined in expressions in the
much the same way as other objects.  The only difference
is that the delimitations is done using curly braces
instead of parenthesis, for example
\begin{verbatim}
{ a + {b + c} }
\end{verbatim}
Here, $a,b,c$ must be parameters, and the expression
is evaluated as soon as it is encountered and replaced
by its equivalent integer value.

All operations which are relevant for integers
are supported (including the ternary operation).
